Probabilistic Catalogs from Crowded Star Fields

### 2012-10-03 paper/ms.tex
comments on the model section

\section{Bayesian Inference}
To quantitatively model uncertainties, Bayesian Inference is the appropriate
framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters
(denoted
collectively by $\theta$) and we expect to obtain some data $x$. Our prior
state of knowledge about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta) \quad.
\end{equation}
\sout{Note that this is a very concise notation \citep{hogg}.}
{\bf I think that I know what you're trying to say here but I think that this
comment is more confusing than helpful...}
We also model how the parameters give rise to the data, via a generative model.
This is also known as a {\it sampling distribution}:
\begin{equation}
p(x|\theta) \quad.
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions over the data space, one for each possible value
of $\theta$. Note that
the sampling distribution is also an assertion of a prior state of knowledge
{\bf (or assumption?)}:
It models prior information about the fact that the data $x$ is connected to
the parameters $\theta$ in some way. Without this prior knowledge, learning is
impossible: there has to be some relationship between the parameters and the
data, otherwise new data tells you about nothing but itself.

@@

\subsection{The Hypothesis Space and the Prior}
Our model makes the following assumptions. There are an unknown number of stars
$N$ in the field. Each star has an unknown
position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
describe the distribution of fluxes (commonly known as the {\it luminosity
function}) of the stars by parameters $\beta$. In summary, the unknown
parameters are:
\begin{eqnarray}
\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N,
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
The prior probability distribution for the unknown parameters can be factorized
using the product rule. With a variety of independence assumptions, the prior
can be written as:
\begin{eqnarray}
p(\theta) = p(N)p(\beta)\prod_{i=1}^N p(x_i, y_i)
p(f_i | \beta)
\end{eqnarray}
Here, we have assumed that the luminosity function does not depend on position.
We have also assumed independence of $N$ and $\beta$, which means that
learning the value of one would not tell us anything about the values of the
others. Finally, the fluxes of the stars come independently from
a common distribution. If we knew the luminosity
function of the stars, then the location and flux of a particular star would
not tell us anything about the location and flux of another star. Really, this
is just a way of implementing exchangeability of the stars.

@@

\subsection{The Sampling Distribution}
The sampling distribution is a probabilistic model for the process that
generates the data; it describes the probability distribution we would use
to predict the data if we happened to know the true catalog. In our case,
the data will be an $n \times n$ array of pixel intensities:
\begin{eqnarray}
\{Y_{ij}\}
\end{eqnarray}
The center of each pixel is:
\begin{eqnarray}
\{x_{ij}, y_{ij}\}.
\end{eqnarray}
Thus, we need
a prescription for simulating an image $\{Y_{ij}\}$ from a catalog $\theta$:
\begin{eqnarray}
\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N,
\left\{f_i\right\}_{i=1}^N\right\}.
\end{eqnarray}
If we knew the true catalog, we could compute the
``mock image'' we would expect to see
in the absence of noise. This mock image (at infinite resolution) is given by:
\begin{eqnarray}
\mathcal{M}(x, y) &=& \sum_{i=1}^N \mathcal{P}(x - x_i, y - y_i)
\end{eqnarray}
where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
Throughout this paper we
assume the pixel-convolved PSF is a known weighted mixture of two concentric
circular
Gaussians with widths $s_1$ and $s_2$:
\begin{eqnarray}
\mathcal{P}(x, y) = \frac{w}{2\pi s_1^2}\exp
\left[
-\frac{1}{2s_1^2}\left(x^2 + y^2\right)
\right]
+ \frac{1-w}{2\pi s_2^2}\exp
\left[-\frac{1}{2s_2^2}\left(x^2 + y^2\right)
\right].
\end{eqnarray}
The finite-resolution observed image is assumed to be generated from the mock
image (evaluated at the center of each pixel) plus Gaussian noise:
\begin{eqnarray}
Y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
\end{eqnarray}
where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed:
\begin{eqnarray}
\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2).
\end{eqnarray}
Throughout this paper, we will assume that the noise variance $\sigma^2$ is
known. This is not usually true in practice (although it is often assumed).
However it is straightforward to extend the model so that $\sigma$ is an
unknown parameter or that the noise variance should be a function of the
intensity of the mock image \citep[e.g.][]{2011MNRAS.412.2521B}.

>>>>

\section{Bayesian Inference}
To quantitatively model uncertainties, Bayesian Inference is the appropriate
framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters
(denoted
collectively by $\theta$) and we expect to obtain some data $x$. Our prior
state of knowledge about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta) \quad.
\end{equation}
\sout{Note that this is a very concise notation} \citep{hogg}.
{\bf I think that I know what you're trying to say here but I think that this
comment is more confusing than helpful...}
We also model how the parameters give rise to the data, via a generative model.
This is also known as a {\it sampling distribution}:
\begin{equation}
p(x|\theta) \quad.
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions over the data space, one for each possible value
of $\theta$. Note that
the sampling distribution is also an assertion of a prior state of knowledge
{\bf (or assumption?)}:
It models prior information about the fact that the data $x$ is connected to
the parameters $\theta$ in some way. Without this prior knowledge, learning is
impossible: there has to be some relationship between the parameters and the
data, otherwise new data tells you about nothing but itself.

@@

{\bf I think that it would be worth commenting on the generative model here
before jumping into the description of the priors\ldots Perhaps something
like:}

The core of our method relies on generating synthetic ``data'' (pixels) given
a set of model parameters (a catalog). This ``generative model'' can be
thought of as an approximation to the propagation of photons from
astronomical objects to the detector. Therefore, it must take the
PSF and other sources of noise into account. In this paper, we will assume
that the pixel-convolved PSF and noise properties are known across the whole
dataset but it would be possible to relax these assumptions with some
non-trivial changes to the method. Since we are especially interested in
the task of generating probabilistic catalogs of crowded stellar fields,
we also assume that the data are fully generated by a collection of only
stars. It wouldn't be very different---if somewhat more complicated---if we
were to also include galaxies and other astronomical sources with analytic
parameterizations of the light profiles in the plane of the sky.

\subsection{The Hypothesis Space and the Prior}
Our model makes the following assumptions. There are an unknown number of stars
$N$ in the field. Each star has an unknown
position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
describe the distribution of fluxes (commonly known as the {\it luminosity
function}) of the stars by parameters $\beta = \{\beta_k \}$.
In summary, the unknown
parameters are:
\begin{eqnarray}
\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N,
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
The prior probability distribution for the unknown parameters can be factorized
using the product rule. With a variety of independence assumptions, the prior
can be written as:
\begin{eqnarray}
p(\theta) = p(N) \, p(\beta) \, \prod_{i=1}^N p(x_i, y_i) \,
p(f_i | \beta)
\end{eqnarray}
Here, we have assumed that the luminosity function does not depend on position.
We have also assumed independence of $N$ and $\beta$, which means that
learning the value of one would not tell us anything about the values of the
other. {\bf $\gets$ Are we making independence assumptions about the individual
$\beta_k$? If so, we should show that in the above equation. Otherwise, this
needs to be singular.} Finally, the fluxes of the stars come independently from
a common distribution. If we knew the luminosity
function of the stars, then the location and flux of a particular star would
not tell us anything about the location and flux of another star. Really, this
is just a way of implementing exchangeability of the stars.

@@

\subsection{The Sampling Distribution}
The sampling distribution is a probabilistic model for the process that
generates the data; it describes the probability distribution we would use
to predict the data if we happened to know the true catalog. In our case,
the data will be an $n \times n$ {\bf (we're already using $N$ for the
number of stars\ldots it might be better to use a different letter here.}
array of pixel intensities:
\begin{eqnarray}
\{Y_{ij}\}
\end{eqnarray}
The center of each pixel is:
\begin{eqnarray}
    \{x_{ij}, y_{ij}\}. \quad \mathrm{couldn't~this~be}\,\{x_i, y_j\}?
\end{eqnarray}
{\bf I also find the notation a little confusing here\ldots I think that we
need different variables for the star positions and the pixel coordinates.}
Thus, we need
a prescription for simulating an image $\{Y_{ij}\}$ from a catalog $\theta$:
\begin{eqnarray}
\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N,
\left\{f_i\right\}_{i=1}^N\right\}.
\end{eqnarray}
If we knew the true catalog, we could compute the
``mock image'' we would expect to see
in the absence of noise. This mock image (at infinite resolution) is given by:
\begin{eqnarray}
\mathcal{M}(x, y) &=& \sum_{i=1}^N \mathcal{P}(x - x_i, y - y_i)
\end{eqnarray}
where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
Throughout this paper we
assume the pixel-convolved PSF is a known weighted mixture of two concentric
circular
Gaussians with widths $s_1$ and $s_2$:
\begin{eqnarray}
\mathcal{P}(x, y) = \frac{w}{2\pi s_1^2}\exp
\left[
-\frac{1}{2s_1^2}\left(x^2 + y^2\right)
\right]
+ \frac{1-w}{2\pi s_2^2}\exp
\left[-\frac{1}{2s_2^2}\left(x^2 + y^2\right)
\right].
\end{eqnarray}
The finite-resolution observed image is assumed to be generated from the mock
image (evaluated at the center of each pixel) plus Gaussian noise:
\begin{eqnarray}
Y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
\end{eqnarray}
where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed:
\begin{eqnarray}
\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2).
\end{eqnarray}
Throughout this paper, we will assume that the noise variance $\sigma^2$ is
known. This is not usually true in practice (although it is often assumed).
However it is straightforward to extend the model so that $\sigma$ is an
unknown parameter or that the noise variance should be a function of the
intensity of the mock image \citep[e.g.][]{2011MNRAS.412.2521B}.
